### 2016-01-22 scripts/figures.Rmd
deciding on legends

---
title: "Figures"
author: "JT McCrone"
date: "December 30, 2015"
output: html_document
---

Here are the figures (or figure panels) we are thinking of using in the Benchmarking paper.  I will include my comments and decriptions on each and perhaps over time this will develop into the results section
```{r,settings_and_lib,echo=F,message=F,}
require(ggplot2)
require("plyr")
require("Biostrings")
require(reshape2)
require("pROC")
require(gridExtra)
require(cowplot)
require(gtable)
require(png)
theme_set(new = theme_classic())
require(knitr)
opts_chunk$set(fig.align = "center", warning = FALSE, echo = FALSE,message=F,cache=F)

rhg_cols<-c("#08519cff","#3182b4ff","#6baed6ff","#b1bdc5ff","#787878ff")

dd=4 # digits in tables
```

### Figure 1
```{r, roc_skematic,fig.height=5,fig.width=3.4}
set.seed(42) # make the figures reproducible
example<-data.frame(Condition=c(rep(x = c(TRUE,FALSE),each=20)),test=c(runif(20,-10,-4),runif(20,-6,0))) # Make a dataframe of T and F positives with random p values that favors a lower distribution for the TP
ex.roc<-roc(example$Condition~example$test) # make an ROC based on this data

cord<-coords(ex.roc,x="all") # get the coordinates of the ROC
cord.l<-melt(cord) # melt the coodinates for ease of plotting
cord.l<- mutate(cord.l,Id=c(0,head(as.numeric(rownames(cord.l))%/%3,-1))) # Add an Id so that the specificity and sensitivity at a given threshold can be groupled together

# Set an id column to group the threshold,specificity, and sensitivity together with the same number by integer division. I adjust with c(0, head ..., -1) since 3%/%3=1 but it should be grouped with the 0's.
roc.df<-dcast(cord.l,Id~Var1) # dcast into columns of threshold, specificity and sensitivity
roc.df<-roc.df[order(roc.df$sensitivity),] # So that the points connect nicely

#Helper function that gets the threshold and sensitivity from and roc at a given speceificity
helper.roc<-function(spec,roc.df){
  above.sense.cut<-subset(roc.df,specificity>=spec,select=c(threshold,specificity,sensitivity)) # get the points where specificity is above the cutoff
  sense.min<-min(above.sense.cut$specificity) # get the minimum of these to get as close to desired sensitivity as possible
  sense.cut<-subset(above.sense.cut,specificity==sense.min) # get all rows where sensitivity is at this point
  spec.max<-max(sense.cut$sensitivity) # but there can be different FPR for each TPR so here we maximize specificity
  output<-subset(sense.cut,sensitivity==spec.max)
}

plot.ex<-function(spec1,spec2,roc.df,distribution){ 
  cut1<-helper.roc(spec1,roc.df)#dataframe with the threshold,sensitivity and specificity at the given specificity
  cut2<-helper.roc(spec2,roc.df)
  output<-rbind(cut1,cut2)
  lines<-data.frame(sense=c(output$sensitivity,output$sensitivity,0,0), # make a data frame with the coordinates for the lines whicha re to be added
                    spec=c(0,0,1-output$specificity,1-output$specificity), # really 1-specificity
                    group=rep(c(spec1,spec2),times=3))

  #plot the roc
  roc.p<-ggplot(roc.df,aes(x=1-specificity,y=sensitivity))+geom_step(color="black",size=1.5)+xlab("1-Specificity")+ylab("Sensitivity")+theme(axis.title.y=element_text(vjust=1))
  roc.p<-roc.p+geom_line(data=lines,aes(x=spec,y=sense,color=as.factor(group)),size=c(1),lty=c(2))+scale_color_brewer(palette="Set1")

 roc.p<-roc.p+theme(axis.title.y = element_text(vjust=1.2))+theme(axis.title.x = element_text(vjust=0))+theme(legend.position="none")

 #plot the points and cut offs
 points.p<-ggplot(distribution,aes(x=1,y=test,fill=Condition))+geom_dotplot(binaxis = "y", stackdir = "center",dotsize = 1.5)+scale_fill_manual(values=c("white","black"),name="",breaks=c(T,F),labels=c("True Positive","False Positive"))+xlab("")+ylab("Log(p value)")+theme(axis.ticks = element_blank(), axis.text.x = element_blank(),axis.title.y = element_text(vjust=1.2),legend.position=c(0.8,0.6))+xlab("")

 points.p<-points.p+geom_hline(yintercept=output$threshold,
                          size=1,lty=2,color=c("#e41a1c","#377eb8")) # these are the first 2 colors in set 1

 plot_grid(points.p, roc.p, labels=c("A", "B"), ncol = 1, nrow = 2)

}
plot.ex(0.35,0.95,roc.df,example)
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/roc.ex.pdf",width = 12,height = 8)
# plot.ex(0.35,0.95,roc.df,example)
# dev.off()

```

```{r,roc_helper_runctions}
find_freq<-function(Id){ # helper function to get the expected frequency from the sample name
  if(Id=="Covaris_5"){
    x=0.05} else if ( Id=="Covaris_25"){
    x=0.025} else if ( Id=="Covaris_125"){
    x=0.0125} else if ( Id=="Covaris_063"){
    x=0.0063} else if ( Id=="Covaris_016"){
    x=0.0016}
}

find_freq.v<-Vectorize(find_freq) # vectorized version of the helper function above

find_freq.h<-function(Id){ # helper function to get the expected frequency from the sample name
  if(Id==0.05){
    x="5.0%"} else if ( Id==0.02){
    x="2.0%"} else if ( Id==0.01){
    x="1.0%"} else if ( Id==0.005){
    x="0.5%"} else if ( Id==0.002){
    x="0.2%"}
}

find_freq.v.h<-Vectorize(find_freq.h) # vectorized version of the helper function above

id<-function(x){
  x<-strsplit(x,".",fixed=T)[[1]][1]
}

seg<-function (x) {
  x <- strsplit(x, ".", fixed = T)[[1]][2]
}

fill_in_plots<-function(x){
  spec<-min(x$adj.specificity[which(x$adj.specificity>0.995)])
  sense<-max(x$adj.sensitivity[x$adj.specificity==spec]) # incase there is a step right here

  y<-data.frame(threshold=x$threshold[1],Id=x$Id[1],adj.specificity=0.9950001,adj.sensitivity=sense,samp=unique(x$samp),exp.freq=x$exp.freq[1],FP=x$FP[1],TP=x$TP[1],sensitivity=x$sensitivity[1],specificity=x$specificity[1])

  if("gc" %in% names(x)){
    y$gc=x$gc[1]
  }
  return(y)
}

sum_roc<-function(x){
  sum.df<-subset(x,category %in% c(TRUE,FALSE),select=c(category,p.val))# get the TRUE and false variant calls and p.vals
  if(length(which(sum.df$category==T))>0){ # filter out cases where there aren't any TP found
    if(length(which(sum.df$category==F))==0){
      sum.df<-rbind(sum.df,data.frame(category=F,p.val=1))
    }  
    roc(sum.df$category~sum.df$p.val,plot=F,CI=T,direction=">")
  }
}

roc_df<-function(roc_analysis){ # get the coordinates and cut offs for all the points in the ROC object
  roc_analysis<-roc_analysis[unlist(lapply(roc_analysis,is.null))==F]
  all<-lapply(roc_analysis,coords,x="all")
  all.long<-lapply(all, melt) 
  all.long<-lapply(all.long, function(x){ 
    mutate(x, 
           Id=c(0,head(as.numeric(rownames(x))%/%3,-1))
    )
  }) # Set an id column to group the threshold,specificity, and sensitivity together with the same number by integer division. I adjust with c(0, head ..., -1) since 3%/%3=1 but it should be grouped with the 0's.
  roc.ls<-lapply(all.long,function(x) dcast(x,Id~Var1)) # dcast into columns of threshold, specificity and sensitivity
  roc.df<-do.call(rbind,roc.ls) # combine into data frame
  roc.df$samp<-unlist(lapply(rownames(roc.df),id)) # add sample id column

  mutate(roc.df, exp.freq=find_freq.v(samp))-> roc.df
}

adjust.coords<-function(roc.df,sum.df,possible_tp,possible_vars){ # adjust the sensitivity and specificity called by pROC  
  samp<-roc.df$samp[1]
  samp.df<-subset(sum.df,Id==samp)
  sense.factor<-length(which(samp.df$category==T))/possible_tp
  TN.samp<-length(which(samp.df$category==F))
  mutate(roc.df,adj.sensitivity=sensitivity*sense.factor,FP=(TN.samp-TN.samp*specificity),TP=adj.sensitivity*possible_tp,adj.specificity=(possible_vars-possible_tp-FP)/((possible_vars-possible_tp)))
}

roc_df.one<-function(roc_anal,thr){ # get the coordinates and cut offs for all the points in the ROC object
  roc_analysis<-roc_anal[unlist(lapply(roc_anal,is.null))==F]
  all<-lapply(roc_analysis,coords,x=thr,input="thr")

  #roc.ls<-lapply(all.long,function(x) dcast(x,Id~Var1)) # dcast into columns of threshold, specificity and sensitivity
  roc.df<-as.data.frame(do.call(rbind,all)) # combine into data frame
  roc.df$samp<-rownames(roc.df) # add sample id column

  mutate(roc.df, exp.freq=find_freq.v(samp))-> roc.df
} 
######### Miseq roc functions ##########
miseq.roc<-function(sum.df,possible_tp,possible_vars){ # A function to run the roc calculations and adjustments
  roc.ls<-dlply(sum.df,~Id,sum_roc)

  roc.df<-roc_df(roc.ls)

  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)

  roc.df.adj<-mutate(roc.df.adj,exp.freq=as.factor(find_freq.v(samp)))
  #fills<-ddply(roc.df.adj,~samp,fill_in_plots) # take out these lines to not fill in the plots
  #roc.df.adj<-rbind(fills,roc.df.adj) # here too
  roc.df.adj<-roc.df.adj[order(roc.df.adj$adj.sensitivity),]
  roc.df.adj$exp.freq <- factor(roc.df.adj$exp.freq, levels = rev(levels(roc.df.adj$exp.freq)))
  return(roc.df.adj)
}
miseq.roc.table<-function(sum.df,cut.off,possible_tp,possible_vars){
  roc.ls<-dlply(sum.df,~Id,sum_roc)
  roc.df<-roc_df.one(roc.ls,cut.off)
  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)
  roc.df.adj<-roc.df.adj[order(roc.df.adj$exp.freq,decreasing = T),]
  roc.table<-subset(roc.df.adj,select=c(exp.freq,adj.sensitivity,TP,adj.specificity,FP))
  return(roc.table)
}

# The ROC data only includes points there the sensitivity changes.  At this point I have limitted our analysis to variants with p<0.05. So the lines in some plots stop short. For the final draft I will not apply this cut off. Even in that future case this step may be need as the ROC data only includes points where the sensitivity changes.  For now   This function extents the lines by adding a data point at the specificity of 0.002 and appropriate sensitivity. At the most it hurts us as we do not increase sensitivity and it might , but I doubt it. Anyway it is of no concern as we only care about the p=0.01 case which is well contained before this extension is needed

######### Hiseq roc functions ##########
hiseq.roc<-function(sum.df,possible_tp,possible_vars){ # A function to run the roc calculations and adjustments
  roc.ls<-dlply(sum.df,~Id,sum_roc)

  roc.df<-roc_df(roc.ls)

  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)
  roc.df.adj<-roc.df.adj[order(roc.df.adj$adj.sensitivity),]
  mutate(roc.df.adj, gc = as.numeric(sub(pat,"\\1",samp)),
       exp.freq = sub(pat,"\\2",samp)
  )-> roc.df.adj
  fills<-ddply(roc.df.adj,~samp,fill_in_plots) # take out these lines to not fill in the plots
  roc.df.adj<-rbind(fills,roc.df.adj) # here too
  roc.df.adj$exp.freq<-as.character(roc.df.adj$exp.freq) # so that teh calculations below work since it is a factor here
  mutate(roc.df.adj,exp.freq=ifelse(test = grepl("0",exp.freq),yes=as.numeric(exp.freq)/1000,no = as.numeric(exp.freq)/100))->roc.df.adj
  roc.df.adj$exp.freq<-as.factor(roc.df.adj$exp.freq)
  roc.df.adj$exp.freq <- factor(roc.df.adj$exp.freq, levels = rev(levels(roc.df.adj$exp.freq)))
  return(roc.df.adj)
}

hiseq.roc.table<-function(sum.df,cut.off,possible_tp,possible_vars){
  roc.ls<-dlply(sum.df,~Id,sum_roc)
  roc.df<-roc_df.one(roc.ls,cut.off)
  roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df,possible_tp,possible_vars)
mutate(roc.df.adj, gc = as.numeric(sub(pat,"\\1",samp)),
       exp.freq = sub(pat,"\\2",samp)
)-> roc.df.adj

mutate(roc.df.adj,exp.freq=ifelse(test = grepl("0",exp.freq),yes=as.numeric(exp.freq)/1000,no = as.numeric(exp.freq)/100))->roc.df.adj
roc.df.adj<-roc.df.adj[order(roc.df.adj$exp.freq,decreasing = T),]
roc.table<-subset(roc.df.adj,select=c(gc,exp.freq,adj.sensitivity,TP,adj.specificity,FP))
return(roc.table)
}

### tableGrob functions and settings #####
tt1<-ttheme_minimal(colhead=list(fg_params=list(fontface=1,cex=0.75,fontfamily="serif")),core=list(fg_params=list(cex=0.75,fontfamily="serif")))
make.table<-function(df,theme){
  tbl<-tableGrob(df,rows=NULL,theme=theme) # Make table
  tbl$widths <- unit(rep(1/ncol(tbl), ncol(tbl)), "npc")
  tbl<-gtable_add_grob(tbl,
        grobs = grobTree(
          segmentsGrob( # line across the bottom
            x0 = unit(0,"npc"),
            y0 = unit(0,"npc"),
            x1 = unit(1,"npc"),
            y1 = unit(0,"npc"),
            gp = gpar(lwd = 2.0)),
        segmentsGrob( # line across the top
            x0 = unit(0,"npc"),
            y0 = unit(1,"npc"),
            x1 = unit(1,"npc"),
            y1 = unit(1,"npc"),
            gp = gpar(lwd = 2.0))),
        t = 1, b = 1, l = 1, r = ncol(tbl))

  tbl<-gtable_add_grob(tbl,
                     grobs=segmentsGrob(
                       x0 = unit(0,"npc"),
            y0 = unit(0,"npc"),
            x1 = unit(1,"npc"),
            y1 = unit(0,"npc"),
            gp = gpar(lwd = 2.0)),
        t = nrow(tbl), b = nrow(tbl), l = 1, r = ncol(tbl))
}
```

### Figure 2
```{r,Miseq_out_of_the_box,fig.width=7,fig.height=5}

sum.df<-read.csv("../data/process/2014-5-30/one.sided/Variants/all.sum.csv",stringsAsFactors=F,comment.char = '#')

true_snv<-read.csv("../data/reference/PR8_WSN33.csv",comment.char = '#') # read in a csv of the differences in PR8 and WSN33 as determined by sanger
true_snv<-subset(true_snv,Ref!="-" & Allele.1!="-") # remove any indels as we are not concerned with them here
mutate(true_snv,mutant=paste0(Name,"_",Ref,Ref.Pos,Allele.1))->true_snv #make a column that has names the variants as segment_ReferencePositionVariant
true_snv<-rename(true_snv,c("Name"="chr","Ref.Pos"="pos")) # rename the columns for use later
PR8_var<-read.csv("../data/reference/PR8_variants.csv",comment.char = "#") # read in differences between plasmid control on viral PR8 used in dilutions

mutate(PR8_var,mutant=paste0(Name,"_",Ref,Ref.Pos,Allele.1))->PR8_var #make a column that has names the variants as segment_ReferencePositionVariant
PR8_var<-rename(PR8_var,c("Name"="chr","Ref.Pos"="pos")) # rename the columns for use later  
sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants to variant calls
sum.df$category[sum.df$mutation %in% PR8_var$mutant]<-"PR8" # id the variants found in PR8
sum.df<-mutate(sum.df,exp.freq=find_freq.v(Id)) # add the expected frequency of the true variants

# Get information about the regions we can investigate based on the primer sites

regions.bed <-read.csv("../data/reference/pr8_noprimingsites.bed.csv",stringsAsFactors = F,comment.char = "#")
regions.bed<-mutate(regions.bed,length=stop-start-1) #because the numbers in the csv are the inner most binding sites of the primers and represent the last positiions that cannot be interogated.

possible_vars<-sum(regions.bed$length)*3 # three possible variants/ position

# limit the PR8 and the true positive data.frames to those not excluded by the priming sites

primer.cut<-function(x){ # a helper function to do this
  chr<-unique(x$chr)
  start<-regions.bed$start[match(x$chr,regions.bed$chr)]
  stop<-regions.bed$stop[match(x$chr,regions.bed$chr)]

  subset(x,pos>start & pos<stop)
}
sum.df<-ddply(sum.df,~chr,primer.cut) # interogating only the sites within primer location removes 244 called variants for two sided and 406 for one sided test
true_snv<-ddply(true_snv,~chr,primer.cut) # interogating only the sites within primer location removes 8 potential true positive (491)
PR8_var<-ddply(PR8_var,~chr,primer.cut) # interogating only the sites within primer location removes 12 potential PR8 differences (15)

## These functions make an ROC by allowing pROC to calculate the positions based on only the variants that are provided, and then we adjust the senesitivity for with what we know.  For example if only 3 TP are found in one sample pROC will call that 100% Sensitivity but we will adjust it to 3/possible_tp and so forth.

tp<-unique(true_snv$mutant)
possible_tp<-length(tp)

m.roc.df<-miseq.roc(sum.df,possible_tp,possible_vars)

miseq.roc.p<-ggplot(m.roc.df,aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_manual(values=rhg_cols,name="Frequency",labels=c("5.0%","2.5%","1.25%","0.63%","0.016%"))+xlab(bquote("1-Specificity (x"~10^-3~")"))+ylab("Sensitivity")+scale_y_continuous(limits=c(0,1))+scale_x_continuous(limits=c(0,0.005),breaks=c(0,0.001,0.002,0.003,0.004,0.005),labels=c(0:5))+theme(legend.position=c(1,.5))

#miseq.roc.p

# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/miseq_roc.one.sided.pdf")
#  miseq.roc.p
# dev.off()

m.roc.table<-miseq.roc.table(sum.df,0.01,possible_tp,possible_vars)

#knitr::kable(m.roc.table)
m.roc.table<-rename(m.roc.table,c("exp.freq"="Frequency","adj.sensitivity"="Sensitivity","TP"="True Positives","adj.specificity"="Specificity","FP"="False Positives"))

m.roc.table$Frequency<-c("5.0%","2.5%","1.25%","0.63%","0.016%")
m.roc.table$Sensitivity<-round(m.roc.table$Sensitivity,digits=dd)
m.roc.table$Specificity<-round(m.roc.table$Specificity,digits=dd)

tbl<-make.table(m.roc.table,tt1)

readPNG("../results/pictures/miseq_setup.png")->img # read in set up png
g <- rasterGrob(img, interpolate=TRUE)
qplot(1:1, 1:1, geom="blank") +
annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
geom_blank()+xlab("")+ylab("")+
theme(line = element_blank(),
text = element_blank(),
line = element_blank(),
title = element_blank())->miseq.setup # empty plot with setup png a background

#plot_grid(miseq.setup,miseq.roc.p, tbl, labels=c("A", "B","C"), ncol = 3, nrow = 1)
# Figure 2

ggdraw() +
  draw_plot(miseq.setup, 0.01, .1, 0.4, 0.8) + #x y width height
  draw_plot(miseq.roc.p, 0.43, 0.42, 0.5, 0.51) +
  draw_plot(tbl, 0.45, 0, 0.53, 0.47) +
  draw_plot_label(c("A", "B", "C"), c(0.01, 0.42, 0.42), c(0.78, 0.95, 0.46), size = 15)

pdf("../results/figures/miseq1.pdf",width=7,height=5)
ggdraw() +
  draw_plot(miseq.setup, 0.01, .1, 0.4, 0.8) + #x y width height
  draw_plot(miseq.roc.p, 0.43, 0.42, 0.5, 0.51) +
  draw_plot(tbl, 0.45, 0, 0.53, 0.47) +
  draw_plot_label(c("A", "B", "C"), c(0.01, 0.42, 0.42), c(0.78, 0.95, 0.46), size = 15)
dev.off()

#ggplot(subset(sum.df,p.val<0.01),aes(x=pos,y=Read_pos,color=category))+geom_point()+facet_wrap(~chr)+geom_hline(yintercept=c(63,188))

#ggplot(sum.df,aes(x=Read_pos,fill=category))+geom_histogram(position="dodge")

```

```{r,Miseq_read.cut,eval=F}
### With read position cut off 63 and 188
read.df<-subset(sum.df,Read_pos>63 & Read_pos<188)
m.roc.df<-miseq.roc(read.df,possible_tp,possible_vars)

miseq.roc.p<-ggplot(m.roc.df,aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_manual(values=rhg_cols)+xlab("1-Specificity")+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1))#+scale_x_continuous(limits=c(0,0.005)

miseq.roc.p

# pdf("../results/figures/miseq_roc.one.sided.pdf")
# miseq.roc
# dev.off()

m.roc.table<-miseq.roc.table(read.df,0.01,possible_tp,possible_vars)
knitr::kable(m.roc.table)

ggplot(subset(sum.df,p.val<0.01),aes(x=Phred,y=MapQ,color=category))+geom_point(alpha=1)

rm(list = setdiff(ls(), lsf.str())) # remove all variables except the functions
tt1<-ttheme_minimal(colhead=list(fg_params=list(fontface=1,cex=0.75,fontfamily="serif")),core=list(fg_params=list(cex=0.75,fontfamily="serif")))

```

### Figure 3

```{r,hiseq_roc,fig.width=7,fig.height=5}
sum.df<-read.csv("../data/process/2015-6-23/Variants/all.sum.csv",stringsAsFactors = F,comment.char = "#") # reading in the variants from the pipeline output

# Getting genome lenght and segment information from the wsn33 fasta file
reference.fasta<-"../data/reference/wsn33_wt_plasmid.fa" 
segments <- fasta.seqlengths(reference.fasta)
regions.bed <- data.frame(chr = gsub("[ ].*","", names(segments)), start=12, stop=segments-13, row.names=NULL) # the univeral primers are 12 and 13 bp long
regions.bed<-mutate(regions.bed,chr=as.character(chr))

# The samples are named in the following format log(genome copy/ul)_expected frequency of variants We'll use that information here to get the gc# and exp.freq for each variant called

pat<-"([0-9]+)_([0-9]+)" 
mutate(sum.df, gc = sub(pat,"\\1",Id),
       exp.freq = sub(pat,"\\2",Id)
)-> sum.df
mutate(sum.df,exp.freq=ifelse(test = grepl("0",exp.freq),yes=as.numeric(exp.freq)/1000,no = as.numeric(exp.freq)/100))->sum.df

sum.df<-subset(sum.df,exp.freq>0.001) # remove this to include exp.freq 0.01%
# Id the variants as T or F 
true_snv<-read.csv("../data/reference/mutant_id.csv",stringsAsFactor=F) # get the T variants
sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants
wt1_mut<-subset(sum.df,Id=="WT1",select=mutation)
wt2_mut<-subset(sum.df,Id=="WT2",select=mutation)

wt_mut<-intersect(wt1_mut$mutation,wt2_mut$mutation)

length(wt_mut)
sum.df$category[sum.df$mutation %in% wt_mut]<-"wt"
sum.df<-subset(sum.df,!(grepl("WT",Id))) # remove the WT samples

possible_tp<-20 # there are 20 possible true positives in the mix
regions.bed<-mutate(regions.bed,length=stop-start-1) #because the numbers in the csv are the inner most binding sites of the primers and represent the last positiions that cannot be interogated.

possible_vars<-sum(regions.bed$length)*3 # three possible variants/ position

h.roc.df<-hiseq.roc(sum.df,possible_tp,possible_vars)

x<-ddply(h.roc.df,~Id,fill_in_plots)

x<-mutate(x,frequency=find_freq.v.h(exp.freq))

Hiseq.roc<-ggplot(subset(h.roc.df,gc==5),aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_manual(values=rhg_cols,name="Frequency",labels=c("5.0%","2.0%","1.0%","0.5%","0.2%"))+xlab(bquote("1-Specificity (x"~10^-3~")"))+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1.1))+scale_x_continuous(limits=c(0,0.005),breaks=c(0,0.001,0.002,0.003,0.004,0.005),labels=c(0:5))+theme(legend.position="none")+geom_text(data=subset(x,gc==5),aes(x=1-adj.specificity,y=adj.sensitivity,label=frequency),hjust=-0.2,position=position_jitter(h=0.1,w=0))

# This is so that the text is not cut off
Hiseq.roc <- ggplot_gtable(ggplot_build(Hiseq.roc))
Hiseq.roc$layout$clip[Hiseq.roc$layout$name == "panel"] <- "off"

h.table.df<-hiseq.roc.table(sum.df,0.01,possible_tp,possible_vars)

#knitr::kable(subset(h.table.df,gc==5))

h.table.df<-rename(h.table.df,c("exp.freq"="Frequency","adj.sensitivity"="Sensitivity","TP"="True Positives","adj.specificity"="Specificity","FP"="False Positives"))

h.table.df<-subset(h.table.df,gc==5,select=-c(gc))
h.table.df$Frequency<-c("5.0%","2.0%","1.0%","0.5%","0.2%")
h.table.df$Sensitivity<-round(h.table.df$Sensitivity,digits=dd)
h.table.df$Specificity<-round(h.table.df$Specificity,digits=dd)

tbl<-make.table(h.table.df,tt1)    # Make table

readPNG("../results/pictures/hiseq_setup.png")->img # read in set up png
g <- rasterGrob(img, interpolate=TRUE)
qplot(1:1, 1:1, geom="blank") +
annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
geom_blank()+xlab("")+ylab("")+
theme(line = element_blank(),
text = element_blank(),
line = element_blank(),
title = element_blank())->hiseq.setup

ggdraw() +
  draw_plot(hiseq.setup, 0, .1, 0.4, 0.8) + #x y width height
  draw_plot(Hiseq.roc, 0.43, 0.43, 0.5, 0.5) +
  draw_plot(tbl, 0.45, 0, 0.53, 0.47) +
  draw_plot_label(c("A", "B", "C"), c(0.05, 0.4, 0.42), c(0.95, 0.95, 0.46), size = 15)

pdf("../results/figures/hiseq1.pdf",width=7,height=5)
ggdraw() +
  draw_plot(hiseq.setup, 0, .1, 0.4, 0.8) + #x y width height
  draw_plot(Hiseq.roc, 0.43, 0.43, 0.5, 0.5) +
  draw_plot(tbl, 0.45, 0, 0.53, 0.47) +
  draw_plot_label(c("A", "B", "C"), c(0.05, 0.4, 0.42), c(0.95, 0.95, 0.46), size = 15)
dev.off()

# 
# knitr::kable(subset(h.table.df,gc==4))
```

```{r, phred_mapq filter}
sig.df<-subset(sum.df,p.val<0.01)

mapq_phred.normal<-ggplot(subset(sig.df,category %in% c(T,F) & gc==5),aes(x=Phred,y=MapQ,color=category))+geom_point(alpha=1)+geom_hline(yintercept=20,linetype=2)+geom_vline(xintercept=30,linetype=2)+scale_x_continuous(limits=c(30,40))+scale_color_hue(name="",labels=c("False posivite","True positive"))+ylab("Mean mapping quality")+xlab("Mean base qualtiy (Phred) ")

#mapq_phred.normal

# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/mapq_phred.normal.pdf")
# mapq_phred.normal
# dev.off()

# mapq_phred.stingy<-ggplot(subset(sig.df,category %in% c(T,F) & gc==5),aes(x=Phred,y=MapQ,color=category))+geom_point(alpha=1)+scale_x_continuous(limits=c(30,40))+geom_hline(yintercept=30)+geom_vline(xintercept=35)+scale_color_hue(name="",labels=c("False positive","True positive"))+ylab("Mapping Quality")+xlab("Base Qualtiy (Phred) ")
# 
# mapq_phred.stingy
# 
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/mapq_phred.stingy.pdf")
# mapq_phred.stingy
# dev.off()

#ggplot(subset(sig.df,gc==5 & category %in% c(T,F)),aes(x=Phred,y=MapQ,color=category))+geom_point()

cut.df<-subset(sum.df,MapQ>30 & Phred>35) # This removes the variants that don't pass

cut.df.all<-sum.df
cut.df.all$p.val[cut.df.all$MapQ<30 | cut.df.all$Phred<35]<-1

# h.cut.df<-hiseq.roc(cut.df,possible_tp,possible_vars)
# 
# mapq_phred.roc<-ggplot(subset(h.cut.df,gc==5),aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_brewer(palette = "Set1",name="Frequency",labels=c("5.0%","2.0%","1.0%","0.5%","0.2%"))+xlab(bquote("1-Specificity (x"~10^-3~")"))+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1))+scale_x_continuous(limits=c(0,0.005),breaks=c(0,0.001,0.002,0.003,0.004,0.005),labels=c(0:5))
# 
# mapq_phred.roc
# 
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/mapq_phred.rocpdf")
# mapq_phred.roc
# dev.off()
# 
# h.cut.table<-hiseq.roc.table(cut.df,0.01,possible_tp,possible_vars)
# 
# knitr::kable(subset(h.cut.table,gc==5))
```

This is still for p<0.01 and 10^5 genomes/ul viral transport media.

### Figure 4

```{r,read_pos}
# ggplot(sig.df,aes(x=pos,y=Read_pos,color=category))+geom_point()+facet_wrap(~chr)
# ggplot(subset(sig.df,gc==5),aes(x=pos,y=Read_pos,color=category))+geom_point()+facet_wrap(~chr)
# 
# ggplot(cut.df,aes(x=pos,y=Read_pos,color=category))+geom_point()+facet_wrap(~chr)
# ggplot(subset(cut.df,gc==5),aes(x=pos,y=Read_pos,color=category))+geom_point()+facet_wrap(~chr)

sig.cut.df<-subset(cut.df,p.val<0.01)

#ggplot(sig.cut.df,aes(x=pos,y=Read_pos,color=category))+geom_point()+facet_wrap(~chr)
# ggplot(subset(sig.cut.df,gc==5),aes(x=pos,y=Read_pos,color=category))+geom_point()+facet_wrap(~chr)

read.pos<-ggplot(subset(sig.cut.df,category %in% c(T,F) & gc==5),aes(x=Read_pos,fill=category))+geom_histogram(position='dodge')+scale_fill_hue(name="",labels=c("False positive","True positive"))+xlab("Average position on read")+ylab("Number of variants")+theme(legend.position=c(.5,.7))
# read.pos
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/read.pos.pdf")
# read.pos
# dev.off()

#cut.read.df<-subset(cut.df,Read_pos>32 & Read_pos<94) # This removes the variants that don't pass

cut.read.df.all<-cut.df.all # this set's the p.val of variants that don't pass the cut to 1
cut.read.df.all$p.val[cut.read.df.all$Read_pos<32 | cut.read.df.all$Read_pos>94]<-1

h.cut.read.df<-hiseq.roc(cut.read.df.all,possible_tp,possible_vars)

read.roc<-ggplot(subset(h.cut.read.df,gc==5),aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_manual(values=rhg_cols,name="Frequency",labels=c("5.0%","2.0%","1.0%","0.5%","0.2%"))+xlab(bquote("1-Specificity (x"~10^-3~")"))+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1))+scale_x_continuous(limits=c(0,0.005),breaks=c(0,0.001,0.002,0.003,0.004,0.005),labels=c(0:5))+theme(legend.position="none")

# read.roc
# 
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/read.roc.pdf")
# read.roc
# dev.off()

h.cut.read.table<-hiseq.roc.table(cut.read.df.all,0.01,possible_tp,possible_vars)
#knitr::kable(subset(h.cut.read.table,gc==5))

h.cut.read.table.5<-subset(h.cut.read.table,gc==5,select=-c(gc))
rename(h.cut.read.table.5,c("exp.freq"="Frequency","adj.sensitivity"="Sensitivity","TP"="True Positives","adj.specificity"="Specificity","FP"="False Positives"))->h.cut.read.table.5
h.cut.read.table.5$Frequency<-c("5.0%","2.0%","1.0%","0.5%","0.2%")
h.cut.read.table.5$Sensitivity<-round(h.cut.read.table.5$Sensitivity,digits=dd)
h.cut.read.table.5$Specificity<-round(h.cut.read.table.5$Specificity,digits=dd)

tbl<-make.table(h.cut.read.table.5,tt1) # Make table

ggdraw() +
  draw_plot(mapq_phred.normal, 0, .5, .5, .5) + 
  draw_plot(read.pos, .5, .5, .5, .5) +
  draw_plot(read.roc, 0, 0, 0.5, .5) +
  draw_plot(tbl, 0.5, 0, 0.5, .5)+
  draw_plot_label(c("A", "B", "C","D"), c(0, 0.5, 0,0.5), c(1, 1, 0.5,0.5), size = 15)

```

### Figure 5

```{r,lofreq,fig.height=5,fig.width=3.4}

lofreq.df<-read.csv("../data/process/2015-6-23/lofreq/all.removed.mapq_vcf.vcf_csv.csv")

lofreq.df$p.val<-lofreq.df$QUAL # to make the roc based on the quality scores given by lofreq

lofreq.roc<-hiseq.roc(cut.df,possible_tp,possible_vars)

ggplot(subset(lofreq.roc,gc==5),aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_manual(values=rhg_cols)+xlab("1-Specificity (adjusted)")+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1))+scale_x_continuous(limits=c(0,0.005),breaks=c(0,0.001,0.002,0.003,0.004,0.005),labels=c(0:5))+theme(legend.position=c(1,.5))->lofreq.roc.p

#lofreq claims no further monkey bussiness is needed to we'll just make a table with all its output
ddply(lofreq.df,~Id,summarize,gc=gc[1],exp.freq=exp.freq[1],TP=length(which(category==T)),FP=length(which(category==F)))->lofreq.table
lofreq.table<-lofreq.table[order(lofreq.table$exp.freq,decreasing = T),]

#knitr::kable(subset(lofreq.table,gc==5,select=c(exp.freq,TP,FP)))
lofreq.table<-subset(lofreq.table,gc==5,select=c(exp.freq,TP,FP))

tbl<-make.table(lofreq.table,tt1) # Make table

# ggdraw() +
#   draw_plot(lofreq.roc.p, 0, .5, 0.5, .5) +
#   draw_plot(tbl, 0.5, 0.5, 0.5, .5) +
#   draw_plot_label(c("A", "B"), c(0, 0.5), c(1, 1), size = 15)

plot_grid(lofreq.roc.p, tbl, labels=c("A", "B"), ncol = 1, nrow = 2)

```

### Figure 6

For p<0.01 with phred, mapq and read position cut off.
```{r, frequency_accuracy}
# Here we'll just be using

## Frequency
cut.read.sig.df<-subset(cut.read.df.all,p.val<0.01)
medians<-ddply(subset(cut.read.sig.df,category==T ),~exp.freq+gc, summarize, median=median(freq.var))

# ac_freq<-ggplot(subset(sum.df,category==T),aes(x=exp.freq,y=freq.var,fill=as.factor(exp.freq)))+facet_wrap(~gc)

#cut.read.sig.df$exp.freq<-as.factor(cut.read.sig.df$exp.freq)
#cut.read.sig.df$exp.freq <- factor(cut.read.sig.df$exp.freq, levels = rev(levels(cut.read.sig.df$exp.freq)))

ac_freq<-ggplot(subset(cut.read.sig.df,category==T & gc=="5"),aes(x=exp.freq,y=freq.var,fill=factor(exp.freq,levels=rev(levels(as.factor(exp.freq))))))

ac_freq<-ac_freq+geom_dotplot(binaxis = "y", stackdir = "center",dotsize=1)+scale_color_manual(values=rhg_cols)

ac_freq<-ac_freq+scale_y_log10(limits=c(0.001,0.1),breaks=c(0.001,0.002,0.005,0.01,0.02,0.05,0.1))#+theme(axis.text.y = element_text(colour = c("red","chartreuse4","black","chocolate2","azure4","blue","black"),face="bold"))
ac_freq<-ac_freq+scale_x_log10(limits=c(0.001,0.1),breaks=c(0.001,0.002,0.005,0.01,0.02,0.050,0.1))#+theme(axis.text.x = element_text(colour = c("red","chartreuse4","black","chocolate2","azure4","blue","black"),face="bold"))
ac_freq<-ac_freq+geom_abline(intercept = 0, slope = 1,linetype=2,size=1)
ac_freq<-ac_freq+xlab("Expected Frequency")+ylab("Measured Frequency")+theme(legend.position="none")

ac_freq<-ac_freq+geom_point(data=subset(medians,gc=="5"),mapping=aes(x=exp.freq,y=median),shape=95,size=30)
ac_freq

pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/freq.pdf")
ac_freq
dev.off()

```

### Figure 7

```{r,roc_lower_inputs4}
roc.4<-ggplot(subset(h.cut.read.df,gc==4),aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_manual(values=rhg_cols,name="Frequency",labels=c("5.0%","2.0%","1.0%","0.5%","0.2%"))+xlab(bquote("1-Specificity (x"~10^-3~")"))+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1))+scale_x_continuous(limits=c(0,0.005),breaks=c(0,0.001,0.002,0.003,0.004,0.005),labels=c(0:5))+theme(legend.position="none")

# roc.4
# 
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/roc.4.pdf")
# roc.4
# dev.off()
# 
# knitr::kable(subset(h.cut.read.table,gc==4))

roc.3<-ggplot(subset(h.cut.read.df,gc==3),aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_manual(values=rhg_cols,name="Frequency",labels=c("5.0%","2.0%","1.0%","0.5%","0.2%"))+xlab(bquote("1-Specificity (x"~10^-3~")"))+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1))+scale_x_continuous(limits=c(0,0.005),breaks=c(0,0.001,0.002,0.003,0.004,0.005),labels=c(0:5))+theme(legend.position="none")

# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/roc.3.pdf")
# roc.3
# dev.off()
# 
# 
# knitr::kable(subset(h.cut.read.table,gc==3))

h.cut.read.table.4<-subset(h.cut.read.table,gc==4,select=-c(gc))
rename(h.cut.read.table.4,c("exp.freq"="Frequency","adj.sensitivity"="Sensitivity","TP"="True Positives","adj.specificity"="Specificity","FP"="False Positives"))->h.cut.read.table.4
h.cut.read.table.4$Frequency<-c("5.0%","2.0%","1.0%","0.5%","0.2%")
h.cut.read.table.4$Sensitivity<-round(h.cut.read.table.4$Sensitivity,digits=dd)
h.cut.read.table.4$Specificity<-round(h.cut.read.table.4$Specificity,digits=dd)

tbl.4<-make.table(h.cut.read.table.4,tt1) # Make table

h.cut.read.table.3<-subset(h.cut.read.table,gc==3,select=-c(gc))
rename(h.cut.read.table.3,c("exp.freq"="Frequency","adj.sensitivity"="Sensitivity","TP"="True Positives","adj.specificity"="Specificity","FP"="False Positives"))->h.cut.read.table.3
h.cut.read.table.3$Frequency<-c("5.0%","2.0%","1.0%")
h.cut.read.table.3$Sensitivity<-round(h.cut.read.table.3$Sensitivity,digits=dd)
h.cut.read.table.3$Specificity<-round(h.cut.read.table.3$Specificity,digits=dd)

tbl.3<-make.table(h.cut.read.table.3,tt1) # Make table

ggdraw() +
  draw_plot(roc.4, 0, .5, .5, .5) +
  draw_plot(tbl.4, .5, .5, .5, .5) +
  draw_plot(roc.3, 0, 0, 0.5, .5) +
  draw_plot(tbl.3, 0.5, 0, 0.5, .5)+
  draw_plot_label(c("A", "B", "C","D"), c(0, 0.5, 0,0.5), c(1, 1, 0.5,0.5), size = 15)

```

### Figure 8

```{r,duplicates}
sum.df<-read.csv("../data/process/2015-11-14/Variants/all.sum.csv",stringsAsFactors = F,comment.char = "#")
#sum.df<-read.csv("~/Box Sync/Lab_work/NGS/Benchmarking/data/2015-11-14/all.sum.csv")
pat<-"([0-9]+)_([0-9]+)"
mutate(sum.df, exp.freq = sub(pat,"\\1",Id),
       dup = as.numeric(sub(pat,"\\2",Id)),
)-> sum.df

sum.df<-subset(sum.df,Id!="PC") # remove the plasmid control that was run
mutate(sum.df,Id=paste(4,exp.freq,sep="_"),exp.freq=ifelse(test = grepl("0",exp.freq),yes=as.numeric(exp.freq)/1000,no = as.numeric(exp.freq)/100))->sum.df

sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants
sum.df$category[sum.df$mutation %in% wt_mut]<-"wt"

cut.df<-subset(sum.df,Phred>35 & MapQ>30 & Read_pos<94 & Read_pos>32)

dups<-ddply(cut.df,~exp.freq+mutation,summarize,freq1=freq.var[1],freq2=freq.var[2],p.val1=p.val[1],p.val2=p.val[2],category=category[1],Id=Id[1])
dups$freq2[is.na(dups$freq2)]<-0
dups$p.val2[is.na(dups$p.val2)]<-1

dups<-mutate(dups,p.val=pmax(p.val1,p.val2))

#dups<-subset(dups,p.val!=1)

dups.roc<-hiseq.roc(dups,possible_tp,possible_vars)

dups.roc.p<-ggplot(dups.roc,aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_manual(values=rhg_cols,name="Frequency",labels=c("5.0%","2.0%","1.0%","0.5%","0.2%"))+xlab(bquote("1-Specificity (x"~10^-3~")"))+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1))+scale_x_continuous(limits=c(0,0.005),breaks=c(0,0.001,0.002,0.003,0.004,0.005),labels=c(0:5))+theme(legend.position="none")

# dups.roc.p
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/dups.roc.pdf")
# dups.roc.p
# dev.off()

dups.table<-hiseq.roc.table(dups,0.01,possible_tp,possible_vars)
#knitr::kable(subset(dups.table))

rename(dups.table,c("exp.freq"="Frequency","adj.sensitivity"="Sensitivity","TP"="True Positives","adj.specificity"="Specificity","FP"="False Positives"))->dups.table
subset(dups.table,select=-c(gc))->dups.table
dups.table$Frequency<-c("5.0%","2.0%","1.0%","0.5%")
dups.table$Sensitivity<-round(dups.table$Sensitivity,digits=dd)
dups.table$Specificity<-round(dups.table$Specificity,digits=dd)

tbl<-make.table(dups.table,tt1) # Make table

ggdraw() +
  draw_plot(dups.roc.p, 0, .5, 0.5, .5) +
  draw_plot(tbl, 0.5, 0.5, 0.5, .5) +
  draw_plot_label(c("A", "B"), c(0, 0.5), c(1, 1), size = 15)

exp.meas.freq<-ddply(cut.df,~exp.freq+mutation+category, function(x){
                                                                                      freq1=x$freq.var[x$dup=="1"]
                                                                                      freq2=x$freq.var[x$dup=="2"]
                                                                                      if(length(freq1)==0) freq1=0
                                                                                      if(length(freq2)==0) freq2=0
                                                                                      out=data.frame(freq1=freq1,freq2=freq2)})

# ggplot(subset(exp.meas.freq),aes(x=freq1,y=freq2,color=category))+geom_point(aes(size=3))+scale_y_log10()+scale_x_log10()+xlab("Frequency in sample 1")+ylab("Frequency in sample 2")+scale_color_hue(name="",labels=c("False positive","True positive"))+theme(legend.position=c(0.5,0.95))+geom_abline(intercept=0,slope=1,lty=2)->dup.freq
# dup.freq
# 
# pdf("~/Box Sync/Lab_work/Talks/2016_812/figures/dup.freq.pdf")
# dup.freq
# dev.off()

```

These are variants that pass mapq,phred, and read cut and p<0.01 in both samples.10^4

This is so much better than before.