### 2016-02-23 doc/paper/results.Rmd

---
output: pdf_document
---

```{r,message=F,echo=FALSE}
xamp_roc<-"figure 1"
miseq<-"figure 2"
Miseq<-"Figure 2"
hiseq<-"figure 3"
Hiseq<-"Figure 3"
qual<-"figure 4"
Qual<-"Figure 4"
freq<-"figure 5"
Freq<-"Figure 5"
lofreq<-"Figure 6"
Lofreq<-"figure 6"
lower<-"figure 7"
Lower<-"Figure 7"
dups<-"figure 8"
Dups<- "Figure 8"
Table<-"Table 1"
table<-"table 1"

potential_var<-"XX"
```

The ability to reliably distinguish true from false single nucleotide variants (SNV) in next generation sequencing data is integral to accurate studies of viral diversity. Sensitivity is the proportion of true variants identified; while, specificity is the proportion of true negatives identified. $1-\text{specificity}$ is the false positive rate, the ratio of false positive variants to true negatives. Both a sensitivity and specificity of 1 describe perfect accuracy in which all the true variants are found and nothing else.

There is an obvious trade off between sensitivity and specificity. Improved sensitivity often requires less stringent criteria in variant calling which reduces specificity. Conversely, increased stringency can improve specificity but reduce sensitivity. This relationship can be represented as a receiver operating characteristic (ROC) curve (`r xamp_roc`). An ROC plots the sensitivity of an assay along the y axis and 1-specificity along the x axis. To make an ROC curve for a variant calling algorithm one must test the assay on known data. The outcomes can then be stratified according to a metric that quantifies the probability that a given variant is real; often this is a p value or quality score. In a controlled benchmakring experiment, all true variants are known, and the sensitivity and specificity can be calculated at different cut offs (`r xamp_roc` A). These points are then used to construct the curve (`r xamp_roc` B). A perfect ROC curve in which all the true positives can be separated from all false positives is a right angle that follows the upper left perimeter of the plot.

#### Initial accuracy

It is beyond the scope of this work to comprehensively compare all variant calling methods employed by virologists. Instead we focus our efforts towards robustly benchmarking one variant caller and highlighting a few means for improving accuracy. In the process we demonstrate the importance of robustly benchmarking any variant calling pipeline under the experimental conditions to which it is being applied.

 We have chosen DeepSNV as our variant caller of choice for this analysis. DeepSNV uses a clonal, plasmid-derived, control to estimate local error rates across the genome. Because it is clonal, the sequence of the control is known with a high degree of confidence and so any nonconsensus base is considered noise. The control is processed together with experimental samples, and so it has identical noise characteristics. DeepSNV applies a hierarchical binomial model at each genomic position and identifies true variants as those with frequencies greater than expected based on the noise found in the plasmid control. Like many variant callers the accuracy of DeepSNV was initially determined using samples that required minimal PCR amplification. We created defined mixtures of two influenza strains, WSN33 and PR8 to characterize the accuracy of DeepSNV when applied to RT-amplified RNA virus populations (`r Miseq` A).

WSN33 and PR8 genomes were mixed serially such that WSN33 was present at relative frequencies of 5, 2.5, 1.25, 0.63 and 0.16%. Because these viruses differ at 491 positions,there are 491 true positives in each dilution. On plasmids subjected to limited amplification, DeepSNV has a sensitivity of 0.860 and perfect specificity for SNV at 0.1% frequency. Under our experimental conditions and with a slightly more stringent p value cutoff ( p=0.01 vs. p=0.05) we found a moderate reduction in sensitivity  (0.830 for variants at 0.63% and 0.14 for variants at 0.16%). The specificity was above 0.998 in all dilutions. While this drop in specificity appears minor, it is important to keep in mind that even a false positive rate of 0.002 results in 78 false positives when applied across the xx,xxx potential variants in the 13.xkb influenza genome.

#### An experimental intrahost population

Although our first benchmarking experiment validated our ability to accurately detect rare variants in influenza populations, the experiment was run under relatively artificial conditions. The virus populations found in patient-derived samples are expected to be less divergent than WSN33 and PR8. Also the nucleic acid concentration in patient samples is much lower than that found samples grown in cell culture. To mimic these conditions, we generated 20 viral colones, each with a single point mutation in the WSN33 background. We sequenced these mutants on the Illumina Miseq platform and confirmed that no additional mutations >1% frequency had arisen between transfection and the passage 1 stock. We also quantified the genome copy number of each stock using quantitative PCR. We then mixed equal genome equivalents of these 20 viruses, to generate an artificial population with $10^5$ copies per microliter and each mutation  present at 5% frequency. This population was serially diluted into a stock of WSN33, generating artificial populations with all 20 mutations present a 2,1,0.5,0.2, and 0.1% frequency (`r Hiseq` A). We then serially diluted each of these populations into basal media to obtain populations with lower input. The range ($10^3-10^5$ copies per microliter) matches the inputs typically found in patient-derived samples (data not shown). We sequenced the samples on the Illumina Hiseq platform, and variants were called using DeepSNV. We also processed and sequenced the WSN33 stock in duplicate to control for any background mutations.

We considered the 20 mutations present in our initial viral mixture as the only true positives in our analysis. Any variant that was present in both duplicates of the wild type stock was masked and  considered neither a true positive nor a true negative. This was done to avoid benchmarking our pipeline with variants identified by our pipeline.

The samples with the greatest input concentration and variant frequencies of 1% and greater had the highest sensitivity (`r hiseq` B and C). Sensitivity dropped at the 0.5% frequency, and in general was lower than that seen previously in the cell culture samples. Our cell culture samples had an  input concentration of at least $10^8$ copies per microliter; whereas, these samples had $10^5$ copies per microliter. The reduced sensitivity is likely due to the fact that there were only 7,000 of each variant genome in the 0.5% frequency samples compared to the equivalent of 62,500,000 in the 0.63% sample from the cell culture experiment.

> This is a little tricky since we amplified the genomes first and then mixed them.

During library prep, a disproportionate amount of the 2.0% sample was lost compared to the others, which likely accounts for the reduced sensitivity relative to the 1.0% sample.

The specificity was greater than 0.99 in all the samples with 10^5^ genomes per microliter. The decreased input concentration and the accompanying increase in RT-PCR cycles is likely the cause of reduced specificity compared with that seen in the cell culture experiment. As above, while 0.99 specificity appears robust, a false positive rate of 0.01 results in over 200 false positive variants when applied to the `r potential_var` potential variants in the influenza genome. Increased specificity can be achieved by applying a more stringent p-value cutoff. However, as shown by the ROC curves in `r hiseq`, this move towards the y axis would markedly reduce sensitivity. Our data demonstrate that with moderate concentrations of input nucleic acid, even statistically significant p-values from a robust variant caller are not sufficient to accurately separate true from false positive variants.

#### Additional filtering criterea

Many next generation sequencing studies utilize mapping (MapQ) and base (Phred) quality cut-offs to ensure that only the highest caliber sequencing data is used to call variants. Mapping quality measures the probability that a given read is mapped to the correct position in the genome while base quality estimates the likelihood that the base call by the sequencer is correct. In our initial analysis we masked bases that had a base quality score less than 30 and did not apply any mapping quality cut offs. We therefore sought to determine if we could improve our specificity by imposing more stringent MapQ and Phred cut-offs. Initially,frequently employed cut-offs such as a MapQ of 20 and a Phred of 30 but were unable to distinguish true from false positives in our 10^5^ samples (`r Qual` A). Many false positives occur on well mapped reads with high quality base calls.

We further parsed our false variant calls by locating them within individual sequencing reads. It is well known that sequence quality drops near the end of a read, and we found that our false positives clustered to these regions (`r Qual` B). The average Phred score of these false positives was xx, and no base calls with a Phred less than 30 was included in the analysis, further demonstrating that filtering on quality score alone is insufficient. In contrast, true positives were uniformly distributed across the reads resulting in an average read position near the middle of the read.

Based on these data we applied a number of empirically determined cut-offs and drastically increased our specificity without sacrificing sensitivity (`r Qual` C and D). For a variant to be considered in our analysis we required a mean mapping quality of 30, a mean phred score of 35, and an average read position within the middle 50% of the read. In addition to these quality cut-offs we applied a number of other criteria in an attempt to increase accuracy such as correcting p-values using the Benjamin Hochman p-value correction, utilizing more stringent p values or frequency cut-offs, retaining PCR duplicate reads, trimming the ends of the influenza genome, and employing various distributions to estimate the error rate in the control sample. All of these results have be summarized in an interactive shiny application available for download at https://github.com/lauringlab/benchmarking_shiny.git and we invite the reader to explore the affects of all these criteria on our data set.

DeepSNV was able to accurately measure the frequency of the true positive variants (`r Freq`). However, while the median frequencies are near the expected values, there is a fair spread in each sample (mean standard deviation of xx). This error should  be kept in mind when employing down stream analysis that depend on frequency measurements such as diversity metrics and comparisons of population structure.

#### Alternative variant callers

DeepSNV is one of many variant callers that use a combination of empiric and statistical approaches to model error rates and separate true from false variants. We asked whether the decreased accuracy observed in our data set was due simply topeculiarities specific to the DeepSNV algorithm. We therefore analyzed our $10^5$ input populations using Lofreq, another alogorthim that is commonly used in viral next generation sequencing studies. Lofreq has been reported to have perfect specificity in the past. Under our conditions Lofreq had marginally reduced sensitivity compared DeepSNV when applied to the variant frequencies \geq 1.0% but marginally increased sensitivity when applied to variant frequencies \l 1.0% (`r Lofreq`). The specificity of LoFreq was comparable to what we observed with DeepSNV in our high-input cell culture-derived populations (`r Miseq`), and better than DeepSNV in our initial implementation (`r Hiseq` prior to Phred, MapQ, and read position filtering). This increased specificity is most likely due to the fast that  the LoFreq algorithm takes MapQ and Phred scores into account when calling variants and has a stringent strand bias filter that removes many of the variants fount only at one end of the reads. However even with these additional characteristics, the specificity of LoFreq was lower than our improved DeepSNV pipeline (`r Qual`) with over 40 false positives per sample. It appears that higher than expected false positive rates are not specific to DeepSNV and most likely plague all variant callers applied to patient-derived viral samples.

\subsection{Lower input levels}

Host-derived viral populations vary in copy number and titer by several orders of magnitude. This variability can be attributed to a variety of factors including but not limited to: collection site, ease of of nucleic acid isolation, library preparation, and host and viral variability. To ensure accuracy across a range of input levels, we diluted our experimental populations serially into basal media and applied our emperically determined quality cut-offs to our list of putative variants (`r Lower`). As expected, the sensitivity for minority variants was lower in populations with lower genome copy number. A variant with a 0.5% frequency in a 10^4^ genomes per microliter sample is only present on 700 genomes and any number of these could be lost in the amplification and library preparation process. We also found that specificity was reduced in samples with lower starting copy numbers. This is presumably due to a greater dependence on RT-PCR amplification. These data highlight the importance of controlling for input levels when comparing diversity across experimental samples.

In most cases, RT-PCRs error should be randomly distributed across the amplified region. If RT-PCR errors are responsible for the reduced specificity found at lower input levels, and RT-PCR are randomly distributed across the genome, we would expect such error to be eliminated by processing the samples in duplicate. To test this hypothesis, we performed duplicate RT-PCR reactions on for the 5,2,1, and 0.5% samples with 10^4^ genomes per microliter. This was done using a separate RNA prep to that shown in `r lower` A. The duplicates were processed separately, but sequenced on the same lane of an Illumina Hiseq. We applied the previously mentioned quality cut-offs and required that a given variant be found in both duplicates. This drastically improved our specificity and resulted in an accuracy that was comparable to that found in the 10^5^ samples (`r Dups`).